---
title: "FINAL EXAM"
output: html_notebook
---


```{r}
turtle <- read.csv('turtle.csv')
head(turtle)
```

(a) A group of 30 turtles are moving along the seashore.  It is estimated that, on average, 70% of them will go into the ocean.  What is the probability that exactly 20 turtles go into the ocean?
```{r}
#dbinom(20, 0.70*30)
dbinom(20,30,0.7)
  

```
(b) While running along a road, a person sees turtles on a random basis.  The person sees on average 1 turtle every 5 minutes.  If they run along the road for 25 minutes, what is the probability that they see at least 4 turtles?
```{r}
1 - ppois(4, 0.20*25)
```

(a) The size of turtles (measuring the length of the shell) from end to end is normally distributed, with mean 25 cm and standard deviation 5 cm.  Picking 1 turtle from random, what is the probability that the length of the turtle shell is less than 22 cm?
```{r}
pnorm(22,25,5)

```
(b) I am waiting for my pet tortoise to come walk to me.  The waiting time is exponentially distributed with mean value 8 minutes.  What is the probability that I will wait between 7 and 10 minutes?
```{r}
pexp(10,1/8) - pexp(7,1/8)

```



Among a group of turtles, 70% of them are box turtles.  40% of the turtles overall have yellow markings.  35% of the turtles are both box turtles and have yellow markings.  What percent of the turtles are neither box turtles nor have yellow markings?



Among a group of turtles, most are friendly.  Some of the turtles are snapping turtles. 95% of the group are friendly turtles that are not snapping turtles.  4% of the turtles are snapping turtles which are unfriendly.  1% of all the turtles are friendly snapping turtles. 
(a) What is the probability that a turtle is friendly, given it is a snapping turtle? 


(b) What is the probability that a turtle is friendly, given it is not a snapping turtle



A turtle needs lettuce to eat.  To get this lettuce, she has two options.  She can raid Farm A, where there is a 40% chance of getting 400 grams of lettuce, a 40% chance of getting 260 grams of lettuce, and a 20% of getting no lettuce.  Or she can raid Farm B, where the lettuce she can get is normally distributed with mean value 330 grams and standard deviation 150 grams

(a) Which choice has the better expected value of the amount of lettuce?
```{r}
0.40*400 + 0.40*260 + 0.20*0
```

(b) If the turtle needs 250 grams of lettuce to avoid being hungry, which option maximizes her chances?
```{r}
dbinom(250,330,40/100)
```




A set of 12 box turtles have their shells measured.  The measurements yield a mean value of 14 cm with a standard deviation of 1.5 cm.  If the length of the shells is normally distributed, construct a 99% 2-sided confidence interval for the mean length of the box turtle shell
```{r}
14 + 1.5/sqrt(length(12))*qt(c(0.005,0.995), 12-1)
```



Take a look at the y field in the turtle dataset, which represents the length of the shells for a set of turtles.  For a certain turtle type, the average shell length is 50 cm.  Construct a hypothesis test at the 90% level to determine whether the average shell length in the dataset is different from 50 cm
```{r}
#turtle$y
#sd(turtle$y)

mean(turtle$y) - sd(turtle$y)/sqrt(length(turtle$y)) * qt(0.90,length(turtle$y)-1)

```



There is a particular type of lettuce which some turtles like and other turtles don't.  Scientists don't know what the probability, so they approximate their prior with a grid from 5% to 95% (in steps of 10%) with each of the 10 possibilities having equal probability.  After feeding the lettuce to 15 turtles, 10 turtles like it.  What is the posterior probability that 65% of turtles like the lettuce?
```{r}
theta <- seq(0.5,.95, by = 0.10)
prior <- c(rep(1/length(theta)), length(theta))

# play 12, win 4
n <- 15
wins <- 10 # turtles that like is a win
posterior <- prior*dbinom(wins,n,theta)/sum(prior*dbinom(wins,n,theta))
posterior
```

For the previous problem, suppose instead that the scientists' prior is a Beta(5,3) distribution.  Again, of the 15 turtles that the lettuce is fed to, 10 turtles like it.  What is the posterior probability that less than 50% of turtles like the lettuce?
```{r}
beta(5,3)
beta(15,7)

sim <- rexp(1000000, 1-0.141444) 
mean(sim)
sd(sim) 
```



A turtle is wondering between 3 parts of its habitat: a cave, a rocky area, and a wooded area.  If the turtle is in the cave, in any given hour, there is a 70% chance the turtle stating in the cave, a 20% chance it goes to the rocky area, and 10% chance it goes to the wooded area.  If the turtle is in the wooded area, in any given hour, there is a 50% chance it stays there, a 30% chance it goes to the cave, and a 20% chance it goes to the rocky area.  If the turtle is in the rocky area, there is a 10% chance it stays there, there is 60% chance it does to the wooded area, and a 30% chance it goes to the cave.  If the turtle starts in the cave, what is the probability that, after a long period of time, the turtle is in the cave?
```{r}
P <- cbind(c(0.7,0.2,0.1),c(0.3,0.5,0.2), c(0.3,0.6,0.1))
x0 <- as.matrix(c(1,0,0)) 
x1 <- P %*% x0

eigen(P)$vectors[,1]/sum(eigen(P)$vectors[,1])
sum(eigen(P)$vectors[,1]/sum(eigen(P)$vectors[,1]))

```




Fill in the blanks.
t is sample mean/ std error 
st error of food would be 
What is the expected length for a male turtle that has eaten 100 grams a day while growing?
```{r}
E equation
```
What is the probability that a female turtle that eats 120 grams a day while growing will be at least 25 cm long?
```{r}

```


```{r}
realestate <- read.csv('Real estate.csv')
head(realestate)
```
Using the real estate dataset, produce a linear regression model to predict house price per unit area, using the 2 other variables which correlate the most strongly to it.  Determine whether the model is a good model in terms of the coefficients and by making a suitable comparison to a baseline model.
```{r}
cor(realestate)[8,]
rs_model <- lm(data = realestate, Y.house.price.of.unit.area ~ X1.transaction.date + X3.distance.to.the.nearest.MRT.station)
#rs_model

summary(rs_model)
```

The coefficients are significant, from looking at the t values and Pr(>|t|). The R-squared value and errors
decreased significantly, so this overall is a good model




A regression is performed at lag 2.  The coefficient of the lagged x is 0.451 and the intercept is 8.323.  Below is a table of the most recent values.  Use the model to make prediction for y for January 2023.
```{r}
# lagged regression
summary(lm(31.812~lag(41.581,6)))
summary(lm(31.812~lag(41.581,6)+lag(26.913,7)))

```





A certain random variable X has the form of a binomial distribution with n = 20 and p = 0.4.  Use a simulation to estimate the expected value of X
```{r}
mean(rbinom(100000,20,0.4))
```





