---
title: "quiz"
output: html_notebook
---

```{r}
customers <- rpois(1000000,2*5)
var(customers)
mean(customers)
```



A probability distribution has cdf F(x) = 0.125x3, where x is between 0 and 2.  Invert this to find the quantile function, and create a simulation of 1,000,000 outcomes to estimate the mean value

```{r}
q_2 <- runif(1000000)/32
mean(q_2)
```



Suppose you are taking a ridiculously long final exam, and you will not finish.  You have 2 potential strategies: finish as many questions as you can, or work to get as many possible correct.  If you work quickly, you can do 30 questions, with a 40% chance of getting any question correct.  If you work on getting them correct, you can do 18 questions, but you have a 70% chance of getting any question correct.

(a) Which alternative has the better expected value of number of questions correct?
```{r}
30*0.4
18*0.7
```

(b) Suppose you calculated that you need to get at least 16 questions correct to pass.  Which alternative gives you the better chance to get at least 16 questions correct?

```{r}
rnorm(1000000, 16)
```




The iris dataset is available on R without pulling in a library.

(a) Find the covariance and correlation between Sepal.Width and Sepal.Length.
```{r}
cov(iris$Sepal.Length, iris$Sepal.Width)
cor(iris$Sepal.Length, iris$Sepal.Width)

```

(b) Give the command for finding the covariance matrix of the numerical fields for iris.
```{r}
glimpse(iris)

cov(iris)
```






