---
title: "quiz4"
output: html_notebook
---



A haunted mansion has a variable number of ghosts passing through it.  It is thought that the number of ghosts each night fits an exponential distribution.  Over the last week leading up to Halloween, a pair of intrepid ghost watchers observed the following number of ghosts: 4, 1, 3, 7, 2, 1, 2.  Use maximum likelihood estimation to estimate the parameter for the number of ghosts.
```{r}
samp <- c(4,1,3,7,2,1,2)
log_likelyhood <- function(a) length(sample) * log(a) - (a+1) * sum(log(samp))

optimize(log_likelyhood,c(1,100),maximum = TRUE)
```



After the zombie apocalypse, a group of 20 people are barricaded in a medieval castle.  Outside the castles is a horde of zombies.  The 20 people don't agree about the number of zombies; the average number of zombies seen is 150 with standard deviation 40.  If the number of zombies is normally distributed, create a 99% 2-sided confidence interval for the number of zombies outside the castle.
```{r}
LCL = qnorm(0.01, 150, 40/sqrt(20))
UCL = qnorm(0.99, 150, 40/sqrt(20))

150 + 40/sqrt(20) *qnorm(0.01)
150 + 40/sqrt(20) *qnorm(0.99)

```








A group of vampire hunters are crafting wooden stakes that will be strong enough to kill a vampire.  The minimum pressure for a wooden stake to kill a vampire is 250 Megapascals.  A group of wooden stakes are fabricated with identical properties, and 10 are strength-tested.  The pressure measured that they can withstand is 260 Megapascals, with a standard deviation of 10 Megapascals.  They want to be 99% sure that the stake will kill the vampire. 

(a) What kind of confidence interval will you use?

(b) To a 99% confidence, will the wooden stakes kill vampires?
```{r}
#min needed 250
#n = 10
#max = 260 
#sd = 10
#mean = 250
250 + 10/sqrt(10) *qnorm(0.99)

```











