---
title: "Homework 4 POCOCK"
output: html_notebook
---

STT 380
Homework 4
Katie Pocock
Due Thursday, October 27 at 11:59:59pm


1. Load the penguins dataset available in D2L. Produce a pairs plot with ggpairs, colored by
species. Then use the pairs plot to answer the following questions.
```{r}
penguins <- read.csv('Downloads/penguins.csv')
library(GGally)
ggpairs(penguins, mapping=ggplot2::aes(colour = species))
```

a. Which species has the most members in the dataset?
    Adelie
b. Which other variable has the highest correlation with body mass (for the overall
dataset)?
    Flipper length
c. For Adelie, which other variable has the highest correlation with body mass?
    Bill depth
d. For each field, which species have similar values according to the histogram?
    Bill length: chinstrap and gentoo
    Bill depth: chinstrap adn adelie
    Flipper length: chinstrap and adelie
    Body mass: chinstrap and adelie 

2. Take the function f(x, y, z) = x2 – 3x + y4 – 3y + z2 + 10z + cos(xyz). Find the location of the minimum value. Use c(0, 0 ,0) as the initialization location. Try different initialization locations. Can you find different locations for minima? If you see multiple ones, which one seems to be a global minimum?
```{r}
fm <- function(x) x[1]^2 - 3*x[1] + x[2]^4 - 3*x[2] + x[3]^2 + 10*x[3] + cos(x[1]*x[2]*x[3])

print(optim(c(0,0,0),fm))
```
so the location of the minimum value is (0.9586131,0.7348568,-4.8938219)
```{r}
min_vals <- c()
for (i in seq(-10,10,1)){
  for (j in seq(-10,10,1)){
    for (k in seq(-10,10,1)){
      min_vals <- append(min_vals, optim(c(i,j,k),fm)$value)
    }
  }
}
summary(min_vals)
min_val <- min(min_vals)
min_val

```

```{r}
optim(c(10,10,-10),fm)[1:2]

```
so the minimum value of -30.12827 is at point c(10,10,-10) and would be the global minimum within the boundaries of -10 to 10 of x,y,z values.

3. One the interval [1, ∞] a probability distribution is defined as p(x) = /x(+1) . Given the sample {2, 3, 2.5, 3, 1.6, 1.4, 1.3, 1.8, 1.9, 2.4, 4.6}, find the log-likelihood, and then calculate the maximum likelihood estimate for the parameter . (Hint: the likelihood function will look very similar to what we had in ICA 14, #2)
```{r}
samp <- c(2, 3, 2.5, 3, 1.6, 1.4, 1.3, 1.8, 1.9, 2.4, 4.6)
log_likelyhood <- function(a) length(sample) * log(a) - (a+1) * sum(log(samp))

optimize(log_likelyhood,c(1,100),maximum = TRUE)

```
so the max : 1.000045



4. The beta distribution on the interval 0 < x < 1 has the form Where  is the gamma function. The two parameters α and β are positive constants. Fortunately, you do not need to use this messy expression, since R has a function to express the probability density: p(x) = dbeta(x, alpha, beta) 
a. Take the sample data in D2L called beta_samp.csv. Construct the log-likelihood function using dbeta.
```{r}
beta_samp <- read.csv('Downloads/beta_samp.csv')
beta_samp

p_x <- function(para) (-1)*sum(log(dbeta(beta_samp$x, para[1],para[2])))
#?dbeta()
```

b. Use the optim function in R to find the maximum likelihood estimate for the parameters.
```{r}
optim(c(1,1),p_x)
```
So alpha = 2.97645 and beta = 4.94619

c. Graph the beta distribution for the given parameters between 0 < x < 1.
```{r}
beta = seq(0,1,0.01)
plot(beta, dbeta(beta, 2.97645,4.94619), type = 'l')
```



5. With the penguins dataset, construct a 99% two-sided confidence interval for the mean value for body mass with the t-distribution. Then do the same for each species individually.
```{r}
penguin_mass <- penguins$body_mass_g
penguin_mean <- mean(penguin_mass)
penguin_sd <- sd(penguin_mass)
penguin_size <- length(penguin_mass)

LCL <- penguin_mean + penguin_sd/sqrt(length(penguin_mass)) * qt(0.005, penguin_size-1)
UCL <- penguin_mean + penguin_sd/sqrt(length(penguin_mass)) * qt(0.995, penguin_size-1)
print(LCL)
print(UCL)

```

```{r}
adelie_mass <- penguins[penguins$species=="Adelie",]$body_mass_g
adelie_mean <- mean(adelie_mass)
adelie_sd <- sd(adelie_mass)
adelie_size <- length(adelie_mass)

LCL_adelie <- adelie_mean + adelie_sd/sqrt(length(adelie_mass)) * qt(0.005, adelie_size-1)
UCL_adelie <- adelie_mean + adelie_sd/sqrt(length(adelie_mass)) * qt(0.995, adelie_size-1)
print(LCL_adelie)
print(UCL_adelie)

```
```{r}
chinstrap_mass <- penguins[penguins$species=="Chinstrap",]$body_mass_g
chinstrap_mean <- mean(chinstrap_mass)
chinstrap_sd <- sd(chinstrap_mass)
chinstrap_size <- length(chinstrap_mass)

LCL_chinstrap <- chinstrap_mean + chinstrap_sd/sqrt(length(chinstrap_mass)) * qt(0.005, chinstrap_size-1)
UCL_chinstrap <- chinstrap_mean + chinstrap_sd/sqrt(length(chinstrap_mass)) * qt(0.995, chinstrap_size-1)
print(LCL_chinstrap)
print(UCL_chinstrap)
```
```{r}
gentoo_mass <- penguins[penguins$species=="Gentoo",]$body_mass_g
gentoo_mean <- mean(gentoo_mass)
gentoo_sd <- sd(gentoo_mass)
gentoo_size <- length(gentoo_mass)

LCL_gentoo <- gentoo_mean + gentoo_sd/sqrt(length(gentoo_mass)) * qt(0.005, gentoo_size-1)
UCL_gentoo <- gentoo_mean + gentoo_sd/sqrt(length(gentoo_mass)) * qt(0.995, gentoo_size-1)
print(LCL_gentoo)
print(UCL_gentoo)
```







