---
title: "ICA 14 POCOCK"
output: html_notebook
---

For this problem, we will use the mtcars dataset.
```{r}
data('mtcars')
length(mtcars$qsec)
```

1. Examine the qsec field in the dataset.  Calculate its mean and standard deviation.
```{r}
mean(mtcars$qsec)
sd(mtcars$qsec)
```

2. What is the standard error?
```{r}
sd(mtcars$qsec)/sqrt(length(mtcars$qsec))

```

3. Construct a 95% confidence interval for the mean of the qsec field.  Then, construct a 99% 1-sided confidence interval (lower limit).  What would the result be for the 99% confidence interval if you used a normal distribution in place of the t-distribution?
```{r}
LCL <- mean(mtcars$qsec) + sd(mtcars$qsec)/sqrt(length(mtcars$qsec)) *qt(0.025, mean(mtcars$qsec) - 1)
UCL <- mean(mtcars$qsec) + sd(mtcars$qsec)/sqrt(length(mtcars$qsec)) *qt(0.975, mean(mtcars$qsec) - 1)
UCL
LCL
```
```{r}
mean(mtcars$qsec) + sd(mtcars$qsec)/sqrt(length(mtcars$qsec)) * qt(0.01,mean(mtcars$qsec) -1 )
```

```{r}
qnorm(0.01, mean(mtcars$qsec), sd(mtcars$qsec)/sqrt(length(mtcars$qsec)))
```



4. Construct 95% lower-confidence and upper-confidence intervals (1-sided).
```{r}
LCL <- mean(mtcars$qsec) + sd(mtcars$qsec)/sqrt(length(mtcars$qsec)) *qt(0.025, mean(mtcars$qsec) - 1)
UCL <- mean(mtcars$qsec) + sd(mtcars$qsec)/sqrt(length(mtcars$qsec)) *qt(0.975, mean(mtcars$qsec) - 1)
UCL
LCL
```

5.  How large would the sample size have to be if a 99% confidence interval were to be 0.2 seconds wide?
    it would be the size of the population which in this case is equal to 32. 











