---
title: "STT 380 ICA 3"
output: html_notebook
---
A very wobbly 4-sided die is used in a particular game.  For this die, the pmf is given by

What is the cdf for this die?
 - made graph
If the die is rolled twice, what is the probability of the sum being a 4?
  - (0.3)*(0.15) + (0.15)*(0.3) + (0.2)*(0.2) = 0.13

A fair coin is flipped 12 times.
What is the probability that the coin is heads exactly half of the time?
```{r}
dbinom(6,12,0.5)
```
What is the probability that the coin is heads 3 times or less?
```{r}
pbinom(3,12,0.5)
```
Perform a simulation of this coin flip in R.  (1 simulation of 12 coin flips)
```{r}
rbinom(1,12,0.5)
```

In a certain game, a player has a 20% chance of winning.  
If the player plays the game 6 times, what is the probability of winning at least twice?
```{r}
1 - pbinom(1, 6, .20)
```
What is the probability on winning zero out of 6 times?
```{r}
dbinom(0,6,0.20)
```
Perform 1,000 simulations of playing the game 6 times.  How does the proportion in the simulation of getting a zero compare to your answer for b?
```{r}
sum(rbinom(1000, 6, 0.20)==0)/1000
```

