---
title: "ICA 10 POCOCK"
output: html_notebook
---

Define an exponential random variable with parameter λ = 1/3.  Generate 2 sets of simulations of the random variable (call them exp_1 and exp_2) of 10,000 simulations each.  Create a third variable su which is the sum of exp_1 and exp_2.  
```{r}
val = 1/3

exp_1 <- rexp(10000, val)
exp_2 <- rexp(10000, val)

su <- exp_1 + exp_2
```

Verify that
E(su) = E(exp_1) + E(exp_2)
```{r}
print(mean(exp_1) + mean(exp_2))


print(mean(su))
```

Var(su) = Var(exp_1) + Var(exp_2)
```{r}
print(var(exp_1) + var(exp_2))


print(var(su))
```


Create 1,000,000 simulations of a multivariate normal random variable similar to the one you worked with in ICA 9 (set var(X) = var(Y) = 1.5, and cor(X,Y) = -0.8), and call it mvn.  x will be the first column of mvn and y will be the second column (for example, x = mvn[,1]).  Create a new variable z which is the sum of the two outputs from the simulation.  
```{r}
min <- -4
max <- 5
var_x <- 1.5
var_y <- 1.5
cor_xy <- -0.8
mvn <- cbind(c(var_x,cor_xy*sqrt(var_x)*sqrt(var_y)),c(cor_xy*sqrt(var_x)*sqrt(var_y),var_y))
x <- mvn[,1]
y <- mvn[,2]
z <- sum(x,y)
```

Verify from the experimental data that
Cov(x, y) = -1.2
```{r}
cov(x,y)
```

E(z) = E(x) + E(y)
```{r}
print(mean(z))

print(mean(x) + mean(y))

```

Var(z) = Var(x) + Var(y) + 2Cov(x, y)
```{r}
print(var(x) + var(y) + 2*cov(x,y))


print(var(z))

```


For this problem we will look at sample means of an exponential distribution with mean value 3 (parameter is 1/3).  Create 20,000 simulation sets of 10,000 simulations each for this distribution.  Calculate the means of the 20,000 sets and plot a histogram.  Does the histogram look approximately like a normal distribution?
```{r}
param <- 1/3
um<-rexp(10000, param)
round(mean(um))

largev<- replicate(20000,um)
mean(largev)
hist(largev)
```














