---
title: "STT 380 ICA 2 POCOCK"
output: html_notebook
---

Practice with the sample function.
a. Run a command to create 10 samples of picking a integer from 1 to 50.
```{r}
sample(c(1:50), 10, replace = TRUE)
```
b. Run a command to create 10 samples of picking a integer from 1 to 50, with 2 no
numbers being the same.
```{r}
sample(c(1:50), 10, replace = FALSE)
```

c. Create a sample with the prob option for creating 100 samples of a picking either 0 or 1, with 0 having probability 2/3 and 1 having probability 1/3.
```{r}
sample(c(0:1), 100, replace = TRUE, prob=c(0.66,0.33))
```
d. Repeat (c) without a prob option in the command.
```{r}
as.integer(sample(seq(1,3), 100, replace = TRUE)==3)
```


II. For the table in slide 8, find
a. P(Michigan | non-commuter)

12/32
b. P(commuter | non-Michigan)


III. Suppose two events A and B have P(A) = 0.4 and P(B) = 0.6.
i. What is the probability of both A and B if they are independent?
ii. What is the minimum value of P(A and B)
iii. What is the maximum value of P(A and B)?
