---
title: "ICA5"
output: html_document
date: "2022-09-19"
---

Define a probability density function as p(x) = cx on the interval 0 ≤ X ≤ 4.
Determine what c is.
1/8
What is the probability that X < 2?
(1/8)*(2)^2-0= 0.5

Customers are entering a store at a rate of 1 every 4 minutes.  Once a customer enters, find:
The probability that another customer will arrive in the next 90 seconds.
```{r}
pexp(1.5,0.25)
```
The probability that another customer will arrive in the next 90 seconds, given that a customer will arrive in the next 5 minutes.
```{r}
pexp(1.5,0.25)/pexp(5,0.25)
```
The probability that it will take more than 8 minutes for another customer to arrive?
```{r}
1 - pexp(8,0.25)
```



The silicon wafers produced at fa semiconductor plant have a flatness deviation which fit a normal distribution with mean 10.0 μm with standard deviation 2.5 μm.  If the wafer has a deviation greater than 16.7 μm, it is defective.  The automated quality control system can detect if the deviation is greater than 17.0 μm.  
What is the probability that a wafer is defective?
```{r}
1 - pnorm(16.7,10,2.5)
```

What is the probability that the quality control system will detect a defect in any wafer?
```{r}
1 - pnorm(17,10,2.5)
```

What is the probability that a wafer is defective and the quality control system does not detect the defect?
```{r}
(1 - pnorm(16.7,10,2.5)) - (1 - pnorm(17,10,2.5))
```



