---
title: "ICA 9 POCOCK"
output: html_notebook
---

Open the pizza dataset.
```{r}
pizza <- read.csv("pizza.csv")
pizza
```

a. Calculate the covariance and correlation between moisture and cal using cov and cor
```{r}
cor(pizza$mois, pizza$cal)
cov(pizza$mois, pizza$cal)

```

b. Verify the cov value by calculating E(XY) – E(X)E(Y) (will not be exact)
```{r}
mean(pizza$mois*pizza$cal) - (mean(pizza$mois)*mean(pizza$cal))
```

c. Verify the cor value by diving the covariance by the product of standard deviations
```{r}
cov(pizza$mois, pizza$cal)/(sd(pizza$cal)* sd(pizza$mois))
```

d. Examine the entire correlation matrix. Which variable has the strongest correlation with cal?
```{r}
cor(pizza$cal, pizza$mois)
cor(pizza$cal, pizza$prot)
cor(pizza$cal, pizza$fat)
cor(pizza$cal, pizza$ash)
cor(pizza$cal, pizza$sodium)
cor(pizza$cal, pizza$carb)

```
calorie and fat have the strongest correlation.



2. For a multivariate normal definition, define the mean value to be (1, 2), with Var(X) = 2, Var(Y) =1, and the Corr(X, Y) = -0.5.
a. Create a contour plot for this distribution.
```{r}
library(mvtnorm)
min <- -4
max <- 5
var_x <- 2.0
var_y <- 1.0
cor_xy <- -0.5
Sigma <- cbind(c(var_x,cor_xy*sqrt(var_x)*sqrt(var_y)),c(cor_xy*sqrt(var_x)*sqrt(var_y),var_y))
x <- seq(min, max, by = 0.1)
y <- seq(min, max, by = 0.1)
z <- matrix(nrow=length(x), ncol=length(y))
co_df <- data.frame('x', 'y', 'z')
for (i in 1:length(x)){
  for (j in 1:length(y)){
    z[i,j] <- dmvnorm(c(x[i],y[j]),c(1,2),Sigma)
  }
}
contour(x, y, z)

co_df <- data.frame('x' = x, 'y' = y)
df.grad <- expand.grid(x = seq(-4,4, by = 0.1),y = seq(-4,4, by = 0.1))
dens <- cbind(df.grad, z = dmvnorm(df.grad,c(0,0), Sigma))
```

b. Create 1,000 simulations of the distribution and create a scatterplot for those
simulations
```{r}
library(MASS)
?mvrnorm
plot(mvrnorm(1000, mu = c(1,2)))

```







