---
title: "ICA 7 POCOCK"
output: html_notebook
---
Let us take the probability distribution we worked with in the previous 2 ICA’s, p(x) = 1/8 * x,
where x is between 0 and 4. Recall that the cdf is given by F(x) = 1/16 * x2.
a. Calculate (on paper) the inverse of the cdf function.
b. Create a variable un which is 1,000,000 simulations of the uniform variable on the unit
segment.
```{r}
un <-runif(1000000)
```

```{r}
un_4 <- (runif(1000000)^3)/8
```

c. Put this un variable into the inverse cdf, generating a new variable un_4.
d. Plot a histogram of un_4. Does the shape correspond to the pdf of the original
function?
```{r}
hist(un_4)
```

e. Use the simulation to find E(X) and E(X2). Based on these results, what is the variance?
E(X) ~ mean(2*(runif(1000000)^(1/3))) = 1.499694 (true
value is 1.5)
◦ E(X2) ~ mean((2*(runif(1000000)^(1/3)^2) = 2.399207 (true
value is 2.4)
```{r}
mean((runif(1000000)^3)/8)

mean(((runif(1000000)^3)/8)^2)

```


2. You are running a manufacturing plant which can produce 120,000 units per month. You are
told that the expected demand for next month is uncertain; it is normally distributed with mean
100,000 and standard deviation 25,000. Each unit sold generates $20 in profit.
a. What is the expected number of units sold? (no calculation required)
```{r}

```

b. What is the profit associated with the expected number of units sold? ($20 x the answer
in (a))
```{r}
20*100000
```

c. Generate 1,000,000 simulations of demand with rnorm. Store it in a variable called
demand.
```{r}
demand <- rnorm(1000000,100000,25000)

```

d. Since sales are limited at 120,000 determine the actual amount of units sold for the
simulations. Call it sales_vol. You can use the command sales_vol = pmax(demand,
120000)
```{r}
sales_vol = pmax(demand, 120000)
mean(sales_vol)
```

e. Calculate for the simulations the total profit. Call it profit.
```{r}
profit <- rnorm(1000000,100000*20,25000*20)
mean(profit)
```

f. What is the average of the profit? How does it compare to the profit of the average
(answer in (b))? Discuss and explain the difference?
```{r}

```

This result is an important consequence of calculations with random variables. It is common in business
to calculate an expected profit like in (b), but (f) is a better calculation. This result has been called The
Flaw of Averages by Sam Savage.