---
title: "ICA 13 POCOCK"
output: html_notebook
---

1. For the price field in the diamonds dataset
```{r}
library(ggplot2)
library(dplyr)

glimpse(diamonds$price)
```

a) First plot a histogram of the data.  Does it look approximately exponential?
      Yes the graph looks approximately exponential
```{r}
hist(diamonds$price)
```

b) Adapt the LL optimization code to find the maximum likelihood estimate for an exponential distribution for the price data.
```{r}
f <- function (lamb) length(diamonds$price)*log(lamb)-1*lamb*sum(diamonds$price)
optimize(f,c(0,3),maximum=TRUE)
```

2. A sample consists of {0.1, 0.2, 0.5, 0.7, 0.8, 0.9, 0.95}.  It is thought to fit a probability distribution of the form p(x) = (α+1)xα, for 0 ≤ x ≤ 1.  Find the maximum likelihood estimate for the parameter α.
```{r}
samp <- c(0.1, 0.2, 0.5, 0.7, 0.8, 0.9, 0.95)

product <- prod(samp)
summed <- sum(samp)

LL <- function(x) log(product) + length(diamonds$z) * log(x) - summed*x
optimize(LL, c(0, 4), maximum = TRUE)
```


3. From diamonds dataset take the field z.  Adopt the code to find the MLE’s for the parameters of the normal distribution.
```{r}
glimpse(diamonds$z)
product <- prod(diamonds$z)
summed <- sum(diamonds$z)

LL <- function(x) log(product) + length(diamonds$z) * log(x) - summed*x
optimize(LL, c(0, 4), maximum = TRUE)

```








